---
title: "Seasonal patterns"
output:
  pdf_document:
    includes:
      in_header: LatexPreamble.sty
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

```{r preamble, echo=FALSE, include=FALSE}

library(plyr)
library(ggplot2)
library(scales)
library(xtable)
#library(DataCombine) # it loads dplyr!
library(magrittr)
library("tidyr")
library("knitr")
library(lubridate)
library(tsibble)
library(fable)
library(feasts)
library(dplyr)
library(purrr)
library(gridExtra)

options(xtable.include.rownames = F,
        xtable.booktabs = T,
        xtable.caption.placement = "top")

source("General.R")
# source("Forecasting.main.R")

load("Data/Panel.RData")

plot.path = "paper_plot/"

monthName = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

Panel <- 
  Panel %>% 
  mutate(date_yearmon =  as.yearmon(Formatted.Date))


Panel_all <- Panel
Panel <- filter(Panel,date_yearmon <= "2020-2")



```



# Plotting Application Series 

```{r applicationSeries}
# Figure 2
# AG
AppSeries = subset(Panel_all, State == "AG", c("Formatted.Date", "Title.2", "Title.16", "Concurrent"))

AppSeries$Formatted.Date = as.Date(as.yearmon(AppSeries$Formatted.Date))
AppSeries = melt(AppSeries, id = "Formatted.Date")


ThreeSeries.AG = ggplot(AppSeries, aes(x = Formatted.Date, y = value, colour = variable)) + theme_bw() + 
              geom_line(size = 1 ) + 
              geom_vline(xintercept = as.numeric(as.Date(paste0(2001:2022, "-1-1"))), linetype = 3, size = 0.5) +            
              
              scale_color_manual(values = c("red","blue","orange"), labels = c("SSDI", "SSI", "Concurrent")) + 
              scale_x_date(date_breaks = "1 year", labels = date_format("%Y"), limits = as.Date(c("2000-10-1", "2021-5-1"))) + 
              labs(
                   x = NULL, y = "Number of Applications", colour = NULL)+
              theme(legend.justification=c(0,1), legend.position=c(0.01,0.99))
ThreeSeries.AG


plot_3series <- function(State_slc){
  AppSeries = subset(Panel_all, State == State_slc, c("Formatted.Date", "Title.2", "Title.16", "Concurrent"))
  
  AppSeries$Formatted.Date = as.Date(as.yearmon(AppSeries$Formatted.Date))
  AppSeries = melt(AppSeries, id = "Formatted.Date")
  
  
  ThreeSeries = ggplot(AppSeries, aes(x = Formatted.Date, y = value, colour = variable)) + theme_bw() + 
                geom_line(size = 1 ) + 
                geom_vline(xintercept = as.numeric(as.Date(paste0(2001:2022, "-1-1"))), linetype = 3, size = 0.5) +            
                
                scale_color_manual(values = c("red","blue","orange"), labels = c("SSDI", "SSI", "Concurrent")) + 
                scale_x_date(date_breaks = "1 year", labels = date_format("%Y"), limits = as.Date(c("2000-10-1", "2021-5-1"))) + 
                labs(title = State_slc,
                     x = NULL, y = "Number of Applications", colour = NULL)+
                theme(legend.justification=c(0,1), legend.position=c(0.01,0.99))
  #ThreeSeries
  
  
}


threeSeries_large <- 
  map(c("AG", "CA", "TX", "NY", "FL", "PA", "OH", "MI", "IL", "WI", "NC", "MN"), plot_3series)



threeSeries_large 


cairo_pdf(paste0(plot.path, "3series_large.pdf"), width = 10, height = 4.8, onefile = T)
threeSeries_large 
dev.off()

```





# Explore

```{r eval=FALSE, include=FALSE}

panel_ts <- Panel_all

panel_ts %<>% 
  mutate(date_yearmon = yearmonth(date_yearmon)) %>% 
  rename(idx = index)


panel_ts %<>%  
  tsibble(key = State, index = date_yearmon) 


df.test <- 
  panel_ts$u %>% 
  dplyr::select(State, date_yearmon, Title.2, Title.16, Concurrent) %>% 
  mutate(across(c(Title.2, Title.16, Concurrent), as.numeric ))


df.test.AG <- 
  df.test %>% 
  filter(State == "AG",
         date_yearmon <= yearmonth("2020-02"))


fit.AG <- 
  df.test.AG %>% 
  model(m1 = ETS(Title.2  ~ error("M") + trend("Ad") + season("M")),
        m2 = ETS(Title.2  ~ error("M") + trend("Ad") + season("A"))
        )

df.test.AG$



m1 <- fit.AG$m1[[1]]

report(m1)
components(m1) %>% autoplot

m1_decomp <- components(m1)


m1_decomp %>% 
  gg_season(season, labels = "both")

m1_decomp %>% 
  gg_subseries(season)







components(fit.AG) %>% 
  autoplot


fabletools::forecast(fit.AG, h = 14) %>% 
  autoplot()



# Producing forecasts using ETS
fit.AG.DI <- 
  df.test.AG %>% 
  model(ETS(Title.2  ~ error("M") + trend("Ad") + season("M")))


fit.AG.SSI <- 
  df.test.AG %>% 
  model(ETS(Title.16  ~ error("M") + trend("Ad") + season("M")))

fit.AG.concurrent <- 
  df.test.AG %>% 
  model(ETS(Concurrent  ~ error("M") + trend("Ad") + season("M")))


fc.AG.DI         <- fabletools::forecast(fit.AG.DI, h = 14)
fc.AG.SSI        <- fabletools::forecast(fit.AG.SSI, h = 14)
fc.AG.concurrent <- fabletools::forecast(fit.AG.concurrent, h = 14)


fc.AG.DI %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2018-01"))) + 
  geom_line(aes(x = date_yearmon, y = Title.2), 
            data = panel_ts %>% filter(State == "AG", date_yearmon >= yearmonth("2020-03")),
            color = "red") +
  coord_cartesian(ylim = c(0, 100e3)) + 
  labs(title = paste0("AG",": SSDI, ETS forecast vs actual"))


fc.AG.SSI %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2018-01"))) + 
  geom_line(aes(x = date_yearmon, y = Title.16), 
            data = panel_ts %>% filter(State == "AG", date_yearmon >= yearmonth("2020-03")),
            color = "red") + 
  coord_cartesian(ylim = c(0, 100e3)) +
  labs(title = paste0("AG",": SSI, ETS forecast vs actual") )
  

fc.AG.concurrent %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2018-01"))) + 
  geom_line(aes(x = date_yearmon, y = Concurrent), 
            data = panel_ts %>% filter(State == "AG", date_yearmon >= yearmonth("2020-03")),
            color = "red") + 
  coord_cartesian(ylim = c(0, 100e3)) + 
  labs(title = paste0("AG",": concurrent, ETS forecast vs actual") )


```


# Seasonal patterns using models

```{r}

panel_ts <- 
  Panel_all %>% 
  mutate(date_yearmon = yearmonth(date_yearmon)) %>% 
  rename(idx = index) %>%  
  tsibble(key = State, index = date_yearmon) 

panel_ts %>% names

df_model <- 
  panel_ts %>% 
  dplyr::select(State, date_yearmon, Title.2, Title.16, Concurrent, dlunemply_U, unemply_U, uRate_U,
                AllowRate.Title.2, AllowRate.Title.16, AllowRate.Concurrent,
                Allow.Title.2, Allow.Title.16, Allow.Concurrent ) %>% 
  mutate(across(c(-State, -date_yearmon), as.numeric )) %>% 
  filter(date_yearmon <= yearmonth("2020-02")) %>% 
  group_by(State) %>% 
  mutate(dunemply_U = unemply_U - dplyr::lag(unemply_U, 1),
         disbApp_tot = Title.2 + Title.16 + Concurrent) 



## ETS model, national total

df_model_AG <- 
  df_model %>% filter(State == "AG")

ets_fml_AA <- ~ error("A") + trend("Ad") + season("A")
ets_fml_MM <- ~ error("M") + trend("Ad") + season("M")
ets_fml_   <- ~ error("M") + trend("Ad") + season("M")


mod_ets_SSDI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Title.2 ~.)))

mod_ets_SSI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Title.16 ~.)))

mod_ets_conc <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Concurrent ~.)))


mod_ets_allow.SSDI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Allow.Title.2 ~.)))

mod_ets_allow.SSI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Allow.Title.16 ~.)))

mod_ets_allow.conc <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, Allow.Concurrent ~.)))

mod_ets_allowRate.SSDI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, AllowRate.Title.2 ~.)))

mod_ets_allowRate.SSI <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, AllowRate.Title.16 ~.)))

mod_ets_allowRate.conc <- 
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, AllowRate.Concurrent ~.)))



mod_ets_unemply <-  # number of unemployment
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, unemply_U ~.)))

mod_ets_uRate <-    # unemployment rate
  df_model_AG %>% 
  model(mod_ets = ETS(update(ets_fml_MM, uRate_U ~.)))

mod_ets_dunemply <- # level changes in unemployment 
  df_model_AG[-1,] %>% 
  model(mod_ets = ETS(update(ets_fml_AA, dunemply_U ~ .)))

mod_ets_dlunemply <- # % changes in unemployment 
  df_model_AG[-1,] %>% 
   model(mod_ets = ETS(update(ets_fml_AA, dlunemply_U ~ .)))

## Check seasonality

# disability application

# number
components(mod_ets_SSDI) %>% gg_season(season, label = "both")
components(mod_ets_SSI)  %>% gg_season(season, label = "both")
components(mod_ets_conc) %>% gg_season(season, label = "both")

# allowance
components(mod_ets_allow.SSDI) %>% gg_season(season, label = "both")
components(mod_ets_allow.SSI)  %>% gg_season(season, label = "both")
components(mod_ets_allow.conc) %>% gg_season(season, label = "both")

# allowance rate
components(mod_ets_allowRate.SSDI) %>% gg_season(season, label = "both")
components(mod_ets_allowRate.SSI)  %>% gg_season(season, label = "both")
components(mod_ets_allowRate.conc) %>% gg_season(season, label = "both")




# No. unemployment
components(mod_ets_unemply) %>% gg_season(season, label = "both")

# unemployment rate
components(mod_ets_uRate) %>% gg_season(season, label = "both")

# changes in unemployment
components(mod_ets_dunemply) %>% gg_season(season, label = "both")

# change rate of unemployement
components(mod_ets_dlunemply) %>% gg_season(season, label = "both")


## check decompostion
components(mod_ets_SSDI) %>% autoplot()
components(mod_ets_SSI)  %>% autoplot()
components(mod_ets_conc) %>% autoplot()
components(mod_ets_unemply) %>% autoplot()
components(mod_ets_dlunemply) %>% autoplot()
components(mod_ets_uRate) %>% autoplot()


## convert additive seasonality to multiplicative seasonality
# 
components(mod_ets_dunemply) %>% 
  mutate(season_mul = (dunemply_U - remainder) /  (dunemply_U - remainder - season)) %>% 
  gg_season(season_mul, label = "both")


# Try STL decomposition

x <- 
  df_model_AG %>% 
  model(STL(Title.2 ~ trend() + season()))

components(x) %>% autoplot

components(x) %>% gg_season(season_year)


```



# Seasonal patterns using models, all states
```{r}

panel_ts <- 
  Panel_all %>% 
  mutate(date_yearmon = yearmonth(date_yearmon)) %>% 
  rename(idx = index) %>%  
  tsibble(key = State, index = date_yearmon) 

# panel_ts %>% names

df_model <- 
  panel_ts %>% 
  dplyr::select(State, date_yearmon, 
                Title.2, Title.16, Concurrent, 
                AllowRate.Title.2, AllowRate.Title.16, AllowRate.Concurrent,
                Allow.Title.2, Allow.Title.16, Allow.Concurrent,
                dlunemply_U, unemply_U, uRate_U) %>% 
  mutate(across(c(-State, -date_yearmon), as.numeric )) %>% 
  filter(date_yearmon <= yearmonth("2020-02")) %>% 
  group_by(State) %>% 
  mutate(dunemply_U  = unemply_U - dplyr::lag(unemply_U, 1),
         disbApp_tot = Title.2 + Title.16 + Concurrent) 



# Estimate ETS models for all variables and all states

ets_fml_AA <- ~ error("A") + trend("Ad") + season("A")
ets_fml_MM <- ~ error("M") + trend("Ad") + season("M")

df_model %>% rename(y = "Title.2")

mod_ets <- list(
  
  SSDI = df_model %>% model(SSDI = ETS(update(ets_fml_MM, Title.2 ~.))),
  SSI  = df_model %>% model(SSI  = ETS(update(ets_fml_MM, Title.16 ~.))),
  conc = df_model %>% model(conc = ETS(update(ets_fml_MM, Concurrent ~.))),
  
  allow.SSDI = df_model %>% model(allow.SSDI = ETS(update(ets_fml_MM, Allow.Title.2 ~.))),
  allow.SSI  = df_model %>% model(allow.SSI  = ETS(update(ets_fml_MM, Allow.Title.16 ~.))),
  allow.conc = df_model %>% model(allow.conc = ETS(update(ets_fml_MM, Allow.Concurrent ~.))),

  allowRate.SSDI = df_model %>% model(allowRate.SSDI = ETS(update(ets_fml_MM, AllowRate.Title.2 ~.))),
  allowRate.SSI  = df_model %>% model(allowRate.SSI  = ETS(update(ets_fml_MM, AllowRate.Title.16 ~.))),
  allowRate.conc = df_model %>% model(allowRate.conc = ETS(update(ets_fml_MM, AllowRate.Concurrent ~.))),

  # number of unemployment
  unemply = df_model %>% model(unemply = ETS(update(ets_fml_MM, unemply_U ~.))),
  # unemployment rate
  uRate = df_model %>% model(uRate = ETS(update(ets_fml_MM, uRate_U ~.))),
  # level changes in unemployment 
  dunemply = 
    df_model %>%  
    filter(date_yearmon > yearmonth("2000-10")) %>% 
    model(dunemply = ETS(update(ets_fml_AA, dunemply_U ~ .))),
  
  # % changes in unemployment 
  dlunemply = 
      df_model %>% 
      filter(date_yearmon > yearmonth("2000-10")) %>% 
      model(dlunemply = ETS(update(ets_fml_AA, dlunemply_U ~ .)))
)

# Components
comp_ets <- plyr::llply(mod_ets, components)

x <- 
bind_rows(comp_ets)


x %>% filter(.model == "SSDI", State == "AG")


# marrangeGrob(list(fig_DI, fig_SSI, fig_concurrent), ncol = 3, nrow = 1)

```






# Counter factual analysis 

```{r}

# Function for plotting projected vs actual applications for three types, 
# each in one plot

panel_ts <- 
  Panel_all %>% 
  mutate(date_yearmon = yearmonth(date_yearmon)) %>% 
  rename(idx = index) %>%  
  tsibble(key = State, index = date_yearmon) 


df.test <- 
  panel_ts %>% 
  dplyr::select(State, date_yearmon, Title.2, Title.16, Concurrent) %>% 
  mutate(across(c(Title.2, Title.16, Concurrent), as.numeric ))


plot_counterFct <- function(state_slc){

# state_slc <- "AG"
  
df.test.AG <- 
  df.test %>% 
  filter(State == state_slc,
         date_yearmon <= yearmonth("2020-02"))

# fit.AG <- 
#   df.test.AG %>% 
#   model(ETS(Title.2  ~ error("M") + trend("Ad") + season("M"))
#         )
#report(fit.AG)# components(fit.AG) %>% 
# autoplot


fabletools::forecast(fit.AG, h = 14) %>% 
  autoplot()



# Producing forecasts using ETS
fit.AG.DI <- 
  df.test.AG %>% 
  model(ETS(Title.2  ~ error("M") + trend("Ad") + season("M")))


fit.AG.SSI <- 
  df.test.AG %>% 
  model(ETS(Title.16  ~ error("M") + trend("Ad") + season("M")))

fit.AG.concurrent <- 
  df.test.AG %>% 
  model(ETS(Concurrent  ~ error("M") + trend("Ad") + season("M")))


fc.AG.DI         <- fabletools::forecast(fit.AG.DI, h = 14)
fc.AG.SSI        <- fabletools::forecast(fit.AG.SSI, h = 14)
fc.AG.concurrent <- fabletools::forecast(fit.AG.concurrent, h = 14)

max_y <- 1.15* max(c(fc.AG.DI$.mean, fc.AG.SSI$.mean, fc.AG.concurrent$.mean))


fig_DI <- 
fc.AG.DI %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2008-01"))) + 
  geom_line(aes(x = date_yearmon, y = Title.2), 
            data = panel_ts %>% filter(State == state_slc, date_yearmon >= yearmonth("2020-03")),
            color = "red") +
  coord_cartesian(ylim = c(0, max_y)) + 
  labs(title = paste0(state_slc, ": SSDI"),
       subtitle = "ETS forecast vs actual",
       x = NULL) +
  theme(legend.position = "bottom")

fig_SSI <- 
fc.AG.SSI %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2008-01"))) + 
  geom_line(aes(x = date_yearmon, y = Title.16), 
            data = panel_ts %>% filter(State == state_slc, date_yearmon >= yearmonth("2020-03")),
            color = "red") + 
  coord_cartesian(ylim = c(0, max_y)) +
  labs(title = paste0(state_slc, ": SSI"),
       subtitle = "ETS forecast vs actual",
       x = NULL) +
   theme(legend.position = "bottom")
  
fig_concurrent <- 
fc.AG.concurrent %>% 
  autoplot(df.test.AG %>% filter(date_yearmon >= yearmonth("2008-01"))) + 
  geom_line(aes(x = date_yearmon, y = Concurrent), 
            data = panel_ts %>% filter(State == state_slc, date_yearmon >= yearmonth("2020-03")),
            color = "red") + 
  coord_cartesian(ylim = c(0, max_y)) + 
  labs(title = paste0(state_slc, ": Concurrent"),
       subtitle = "ETS forecast vs actual",
       x = NULL) +
   theme(legend.position = "bottom")


marrangeGrob(list(fig_DI, fig_SSI, fig_concurrent), ncol = 3, nrow = 1)

}


fig_AG <- plot_counterFct("AG")
fig_CA <- plot_counterFct("CA")
fig_TX <- plot_counterFct("TX")
fig_NY <- plot_counterFct("NY")
fig_FL <- plot_counterFct("FL")
fig_PA <- plot_counterFct("PA")
fig_NC <- plot_counterFct("NC")
fig_WI <- plot_counterFct("WI")


cairo_pdf(paste0(plot.path, "counterFct1.pdf"), width = 20, height = 5.6, onefile = T)
fig_AG
fig_CA
fig_TX
fig_NY
fig_FL
fig_PA
fig_NC
fig_WI
dev.off()


# ggsave("text.pdf",  width = 10, height = 4, scale = 1.4)

```



# Describing seasonality with decomposition

```{r}
#    

df.test %>% distinct(State)

## seasonal plots:
# Useful in identifying years in which the pattern changes
df.test %>% 
  filter(State =="AG") %>% 
  gg_season(Title.2, labels = "both")


## seasonal subplots
# Useful in indentifying changes within particular seasons
df.test %>% 
  filter(State =="AG") %>% 
  gg_subseries(Title.2)








```







# 1. Description of Seasonality in Applications and Unemployment 



## Seasaon dummies for disability applications and unemployment

```{r function_for_seasonal_dummies}

seasonDummy = function(ts){
# This function runs OLS regression of a selected series on seasonal dummies
# using "lm" function. The coefficients are then tested using NeweWest autocorrelation
# robust standard errors. The test result and adjusted R squared are reported.

# Input
  # ts: A time series (ts) object. 
  
# Output: A list with 4 elements
  # Model: The regression model of the class lm
  # test: The test result with NewyWest standard errors.
  # R2: R squared
  # R2adj: adjusted R squared. 
  
  month = factor(cycle(ts))
  m = lm( ts ~ month - 1 )
  test = coeftest(m, vcov = NeweyWest)
  
  result = list(model = m, test = test, R2 = summary(m)$r.squared, R2adj = summary(m)$adj.r.squared)
  invisible(result)
}

```


```{r Estimating_Seasonal_Dummies}

end_select <- c(2020,2)

variables = c("dlTotal","dlTitle.2", "dlTitle.16", "dlConcurrent", "dlunemply_U")

dummyNW = t(sapply(variables, function(x){ m = seasonDummy(select(x, "AG", Panel))
                                         c(R2 = m$R2, m$model$coef)}))


Dummy.Total = data.frame(t(sapply(states52, function(x){m = seasonDummy(select("dlTotal", x, Panel, end = end_select  ))
                                             c(R2 = m$R2, m$model$coef)})))
Dummy.Title.2 = data.frame(t(sapply(states52, function(x){m = seasonDummy(select("dlTitle.2", x, Panel, end = end_select))
                                             c(R2 = m$R2, m$model$coef)})))
Dummy.Title.16 = data.frame(t(sapply(states52, function(x){m = seasonDummy(select("dlTitle.16", x, Panel, end = end_select))
                                             c(R2 = m$R2, m$model$coef)})))
Dummy.Concurrent = data.frame(t(sapply(states52, function(x){m = seasonDummy(select("dlConcurrent", x, Panel, end = end_select))
                                             c(R2 = m$R2, m$model$coef)})))
Dummy.unemploy   = data.frame(t(sapply(states52, function(x){m = seasonDummy(select("dlunemply_U", x, Panel, end = end_select))
                                             c(R2 = m$R2, m$model$coef)})))

Dummy.Total$State = rownames(Dummy.Total)
Dummy.Title.2$State = rownames(Dummy.Title.2)
Dummy.Title.16$State = rownames(Dummy.Title.16)
Dummy.Concurrent$State = rownames(Dummy.Concurrent)
Dummy.unemploy$State = rownames(Dummy.unemploy)

Dummy.Total = MoveFront(Dummy.Total, c("State"))
Dummy.Title.2 = MoveFront(Dummy.Title.2, c("State"))
Dummy.Title.16 = MoveFront(Dummy.Title.16, c("State"))
Dummy.Concurrent = MoveFront(Dummy.Concurrent, c("State"))
Dummy.unemploy = MoveFront(Dummy.unemploy, c("State"))


Dummy.Title.2 %>% arrange(R2) %>% head
Dummy.Title.16 %>% arrange(R2) %>% head
Dummy.Concurrent %>% arrange(R2) %>% head

#save(Dummy.Total, Dummy.Title.2, Dummy.Title.16, Dummy.Concurrent)

```


```{r Plot_Seasonal_Dummies}

plot.AppUnempLag = function(Data, region, type, Plot = TRUE){
    ## Data must be one of the Dummy.XXXX data frames. 

# Data = Dummy.Title.2
# region = "AG"
# type = "Total"

DATA = data.frame( Application = as.numeric(Data[Data$State == region, paste0("month",1:12)]),
                   Unemployment = as.numeric(Dummy.unemploy[Dummy.unemploy$State == region, paste0("month",1:12)][c(12,1:11)]),
                   month = 1:12)
DATA = melt(DATA, id = "month")

AppUnempLag =  ggplot(DATA, aes(x = as.integer(month), y = value, colour = variable, linetype = variable)) + theme_bw() + 
               geom_line(size = 0.6) + geom_point(size = I(2.5)) +
               scale_color_manual(values=c("red", "blue"), labels = c(paste(type,"Applications"), "Unemployment(1 month Lag)")) + 
               scale_linetype(labels = c(paste(type,"Applications"), "Unemployment(1 month Lag)")) + 
               scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                                          "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
               labs(x = NULL, y = "Coefficients", colour = NULL, linetype = NULL) + 
               theme(legend.justification=c(1,1), legend.position=c(0.99, 0.99))
if(Plot) print(AppUnempLag)
return(AppUnempLag)
}

AppUnempLag = plot.AppUnempLag(Dummy.Title.2, "AG", "Total")
AppUnempLag

# Figure 4
cairo_pdf(paste0(plot.path, "apply&unemplyLag.pdf"), width = 10*0.8, height = 5*0.8)
AppUnempLag
dev.off()

cairo_pdf(paste0(plot.path, "SeasonalDummies_AllStates.pdf"), width = 10, height = 5, onefile = T)
  for(i in states52) print(plot.AppUnempLag(Dummy.Title.2, i, "DI", FALSE) + labs(title = i) + coord_cartesian(ylim=c(-0.2, 0.2)))
dev.off()

```


```{r Boxplot}
# Figure 5

Dummy.all = rbind(cbind(Dummy.Title.2, type = "SSDI"), 
                  cbind(Dummy.Title.16,type = "SSI"), 
                  cbind(Dummy.Concurrent, type = "Concurrent"))
Dummy.all = subset(Dummy.all, State %in% states50)
Dummy.all$R2 = NULL

Dummy.all = melt(Dummy.all, id = c("State", "type"))
BoxplotDummy = ggplot(Dummy.all, aes(x = variable, y = value)) + theme_bw() + facet_wrap( ~ type , ncol = 1) + 
               geom_boxplot(outlier.size = 1)+
               geom_hline(yintercept = 0, linetype = 3, color = "blue", size = I(0.8)) + 
               scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) + 
               labs(x = NULL, y = "Coefficients")
BoxplotDummy

cairo_pdf(paste0(plot.path, "BoxplotDummycoeff.pdf"), width = 8, height = 12, onefile = T)
  BoxplotDummy  
dev.off()


```


```{r Table_of_seasonal_dummies_for_nationawide}
# Table 3

dummyAG.table = rbind(cbind(Category = "SSDI",        subset(Dummy.Title.2,   State == "AG", paste0("month",1:12))),
                      cbind(Category = "SSI",         subset(Dummy.Title.16,  State == "AG", paste0("month",1:12))),
                      cbind(Category = "Concurrent",  subset(Dummy.Concurrent,State == "AG", paste0("month",1:12))),
                      cbind(Category = "Unemployment",subset(Dummy.unemploy,  State == "AG", paste0("month",1:12)))
                      )
names(dummyAG.table) = c("Category", monthName)
dummyAG.table

print(xtable(dummyAG.table, caption = " Coefficients of Seasonal Dummies: Growth Rate of National Aggregate"))

```


```{r Table_of_R^2_for_top_5}
# Table 2

R2.table = merge( subset(Dummy.Title.2,  State %in% c("AG", top5), c("State","R2")),
                  subset(Dummy.Title.16, State %in% c("AG", top5), c("State","R2")), by = "State")
R2.table = merge( R2.table, subset(Dummy.Concurrent, State %in% c("AG", top5), c("State", "R2")), by = "State")

names(R2.table) = c("State", "SSDI", "SSI", "Concurrent")

R2.table$State = factor(R2.table$State, levels = c("AG","CA","TX","FL","NY","PA"))
R2.table = R2.table[order(as.numeric(R2.table$State)), ]


print(xtable(R2.table, caption = "Magnitudes of Seasonality in Disability Applications"))

# For the in-line text
Dummy.Title.2[Dummy.Title.2$State == "ND",]
Dummy.Title.16[Dummy.Title.16$State == "ND",]
Dummy.Concurrent[Dummy.Concurrent$State == "ND",]

```


```{r Table_of_R2_and_seasonal_dummies_for_top5}
# For Table A1-A3 in appendix

Dummy.table = function(Data, caption){
  
  tab = subset(Data, State %in% c("AG", top5))
  names(tab) = c("State","R2", monthName )
  tab$State = factor(tab$State, levels = c("AG","CA","TX","FL","NY","PA"))
  tab = tab[order(as.numeric(tab$State)),]

  xtab = xtable(tab, caption = caption)

  print(xtab) 
}

Dummy.table(Dummy.Title.2, caption = "R2 and coefficients of seasonal dummies: SSDI (only) Applications")
Dummy.table(Dummy.Title.16, caption = "R2 and coefficients of seasonal dummies: SSI (only) Applications")
Dummy.table(Dummy.Concurrent, caption = "R2 and coefficients of seasonal dummies: Concurrent Applications")


```


## Testing for Heterogeneity

```{r PYtest}
source("Modeling.HeteroPYtestSeason.R")

PanelDummy = cbind(Panel, dummy("Month",Panel)) 
PYSeason.Title.2    = PYSeason.test("dlTitle.2", paste0("Month",1:12), 
                                     cbind(Panel, dummy("Month",Panel)), states50)
PYSeason.Title.16   = PYSeason.test("dlTitle.16", paste0("Month",1:12), 
                                     cbind(Panel, dummy("Month",Panel)), states50)
PYSeason.Concurrent = PYSeason.test("dlConcurrent", paste0("Month",1:12), 
                                     cbind(Panel, dummy("Month",Panel)), states50)

PYSeason.test = rbind( cbind(Category = "SSDI", PYSeason.Title.2[3,], PYSeason.Title.2[4,]),
                       cbind(Category = "SSI" , PYSeason.Title.16[3,], PYSeason.Title.16[4,]),
                       cbind(Category = "Concurrent", PYSeason.Concurrent[3,], PYSeason.Concurrent[4,]))
names(PYSeason.test) = c("Category", "Dh", "p-value", "Dt", "p-value")

PYSeason.test

print(xtable(PYSeason.test))


```


## Plotting Application Series 

```{r applicationSeries}
# Figure 2

AppSeries = subset(Panel_all, State == "AG", c("Formatted.Date", "Title.2", "Title.16", "Concurrent"))

AppSeries$Formatted.Date = as.Date(as.yearmon(AppSeries$Formatted.Date))
AppSeries = melt(AppSeries, id = "Formatted.Date")


ThreeSeries = ggplot(AppSeries, aes(x = Formatted.Date, y = value, colour = variable)) + theme_bw() + 
              geom_line(size = 1 ) + 
              geom_vline(xintercept = as.numeric(as.Date(paste0(2001:2022, "-1-1"))), linetype = 3, size = 0.5) +            
              
              scale_color_manual(values = c("red","blue","orange"), labels = c("SSDI", "SSI", "Concurrent")) + 
              scale_x_date(date_breaks = "1 year", labels = date_format("%Y"), limits = as.Date(c("2000-10-1", "2021-5-1"))) + 
              labs(x = NULL, y = "Number of Applications", colour = NULL)+
              theme(legend.justification=c(0,1), legend.position=c(0.01,0.99))
ThreeSeries

cairo_pdf(paste0(plot.path, "3series.pdf"), width = 10, height = 4.8, onefile = T)
ThreeSeries
dev.off()

```


## Frequency Domain Graphs

```{r Coherence_Between_stochastic_Components}
# Check the coherence between the series net of deterministic seasonality

plot.coherence = function(region, category1, category2 = "dlunemply_U", title = ""){
  
  res.apply = ts(seasonDummy(select(category1, region, Panel))$model$res, s = c(2000,11), f = 12)
  res.emply = ts(seasonDummy(select(category2, region, Panel))$model$res, s = c(2000,11), f = 12)
  sRes = spec.pgram(cbind(res.apply, res.emply),
                  #spans = c(5,5),
                  kernel("daniell", c(3,3)),
                  taper = 0, plot = FALSE, fast = F)
  plot(sRes, plot.type = "coh", ci.lty = 2, main = title)
  abline(v = 1:6, lty = 3)
}

plot.coherence("AG","dlTotal", "dlunemply_U", title = "Total")
plot.coherence("AG","dlTitle.2", "dlunemply_U", title = "SSDI")
plot.coherence("AG","dlTitle.16", "dlunemply_U", title = "SSI")
plot.coherence("AG","dlConcurrent", "dlunemply_U", title = "Concurrent")


cairo_pdf(paste0(plot.path, "coherenceNW.pdf"), width = 9, height = 3, onefile = T)
  par(mfrow = c(1,3))
  plot.coherence("NW","dlTitle.2", "dlunemply_U", title = "SSDI")
  plot.coherence("NW","dlTitle.16", "dlunemply_U", title = "SSI")
  plot.coherence("NW","dlConcurrent", "dlunemply_U", title = "Concurrent")
dev.off()



L = 19
alpha = 0.05
fvalue = qf(1- alpha, 2, 2*L-2)
fvalue
C = (fvalue/(L - 1 + fvalue))
C

qf(1-0.001, 2, 34)

# library(seewave)
# category1 = "dlTitle.16"; category2 = "dlunemply_U"; region = "NW"
# res.apply = ts(seasonDummy(select(category1, region, Panel))$model$res, s = c(2000,11), f = 12)
# res.emply = ts(seasonDummy(select(category2, region, Panel))$model$res, s = c(2000,11), f = 12)
# 
# x = cbind(res.apply, lag(res.emply, -1))
# x = window(x, start = c(2000,12), end = c(2014,5))
# x
# coh(x[,1],x[,2])

```


## Seasonal Pattern in Month plots

```{r seasonal_pattern}

monthplot(select("dlTotal", "AG", Panel))
monthplot(select("dlunemply_U", "AG", Panel))

## Do recursive estimation of seasonal dummies. 

```


## Seasonal Pattern in Allowance Rate

To do list with the allowance and allowance rate

Allowance rate
1. Regression on seasonal dummies, by application type and state. Output is dataframe and table 
2. Plot the coefficients of allowance and application on the same plot. By type and state
    - No lag
    - 4th lag

Application 
1. Regression on seasonal dummies, by application type and state. Output is dataframe and table 


 Main findings:
  - Genrerally, the seasonality in allowance rates is much more moderate than that in application. 
  - Suprisingly, the allowance rates have very similar seasonal pattern, although more moderate, to
      the application series. According to SSA report, the determination process averagely take 131 days
      (see Rutledge 2011,p16), which is approximately 4-5 months. Therefore if the peaks in applications do 
      have effect (negative) on allowance rates, the seasonal pattern of allowance should show peaks in
      June ~July(4-5 months after the Feb peak in application), and Dec~Jan(4-5 months after the Aug peak). 
      None of them, however, shows up the the seasonal pattern in allowance rate. 
  - The similarity between the seasonal patterns in application volume and allowanc rates may imply that 
      there are some come common driving factor behind both of them, which are possibly related to some 
      administrative process/issues. If this is the case, the the seasonality in allowance rate does not 
      support the causal linkage between unemployment and application. 
  - Next Step: Try to look at seasonality in application and allowance rate state by state, to see if there
     is any relationship, which may be masked in the national level analysis.


  - Nationalwide series, CA and PA series have steep downward trend in allowance rate, with a short - 
    period of rebound in the middle of the sample. 
  - NY and TX exhibits very different trend in allowance rates compared to other states, showing hump-shaped   
      trend over the sample period.
  - The allowance rates of 3 different types of applications in FL have different trends. 


 Main findings from plots of applications vs allowance rate(4th lag)
 
  - If we look at the contemporary relationship 
     - Applications and allowance rate have very similar pattern in nationalwide series, which makes no sense if the seasonality in application is caused by unemployment. 
     - However, the similarity is much less obvious in a state-by-state comparison. The common peaks in Feburary appear in most states, while only some of the states exhibit common peaks in August. For the top 5 states, allowance rates in CA, TX, NY, PA do not show siginificant increase in August, while FL does.  
     
 - If we look at application and 4th lag of allowance rate, 
    - The negative relationship are more prominent at the begining and the end of the year. 
    - The relationship in CA, TX, FL are more prominent.
    
 - It is still hard to tell which relationship is real. So far the evidence from the allowance rate is still quite weak.


```{r seasonality_of_AllowRate, eval=FALSE, include=FALSE}

Path <- paste0(plot.path,"/Allowance/")
plot.path


# Seasonal plots of national level allowance rates
select("AllowRate.Total", "AG", Panel) %>% monthplot
select("AllowRate.Title.2", "AG", Panel) %>% monthplot
select("AllowRate.Title.16", "AG", Panel) %>% monthplot
select("AllowRate.Concurrent", "AG", Panel) %>% monthplot


cairo_pdf(paste0(Path, "/Allowance_Monthplot.pdf"), width = 10, height = 8, onefile = T)
 par(mfrow = c(2,2))
# Seasonal plots of first differenced allowance rates
select("AllowRate.Total", "AG", Panel) %>% diff %>% monthplot(main = "dTotal")
select("AllowRate.Title.2", "AG", Panel) %>% diff %>% monthplot(main = "dTitle.2")
select("AllowRate.Title.16", "AG", Panel) %>% diff %>% monthplot(main = "dTitle.16")
select("AllowRate.Concurrent", "AG", Panel) %>% diff %>% monthplot(main = "dConcurrent")
dev.off()


# Estimated Spectral plots of first differenced allowance rates
cairo_pdf(paste0(Path, "/Allowance_Spectrum.pdf"), width = 10, height = 8, onefile = T)
 par(mfrow = c(2,2))
select("AllowRate.Total", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Total")
select("AllowRate.Title.2", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Title.2")
select("AllowRate.Title.16", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Title.16")
select("AllowRate.Concurrent", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Concurrent")
dev.off()

i <- "CA"

select("AllowRate.Total", i, Panel)     %>% monthplot
select("AllowRate.Title.2", i, Panel)   %>% monthplot
select("AllowRate.Title.16", i, Panel)  %>% monthplot
select("AllowRate.Concurrent", i, Panel)%>% monthplot

select("AllowRate.Total", i, Panel) %>% diff %>% monthplot
select("AllowRate.Title.2", i, Panel) %>% diff %>% monthplot
select("AllowRate.Title.16", i, Panel) %>% diff %>% monthplot
select("AllowRate.Concurrent", i, Panel) %>% diff %>% monthplot


```

```{r Estimating_Seasonal_dummies_AllowRate}

variables = c("AllowRate.Total","AllowRate.Title.2", "AllowRate.Title.16", "AllowRate.Concurrent")

select("AllowRate.Total", "AG", Panel) %>% diff
# x <- seasonDummy(select("AllowRate.Total", "CA", Panel) %>% diff); x$test[,"Pr(>|t|)"]



dummyNW_Allow = t(sapply(variables, function(x){ m = seasonDummy(select(x, "AG", Panel) %>% diff)
                                                 p.value <- m$test[,"Pr(>|t|)"]; names(p.value) <- paste0(names(p.value), ".p")
                                                 c(R2 = m$R2, m$model$coef, p.value)}))

Reg_allow <- function(x, type){
  m <- seasonDummy(select(type, x, Panel) %>% diff)
  R2 <- m$R2
  Coeff <- m$model$coef
  p.value <- m$test[,"Pr(>|t|)"]
  names(p.value) <- paste0(names(p.value), ".p")
  c(R2 = R2, Coeff, p.value)
}

Dummy_Allow.Total = data.frame(t(sapply(states52, Reg_allow, type = "AllowRate.Total" )))
Dummy_Allow.Title.2 = data.frame(t(sapply(states52, Reg_allow, type = "AllowRate.Title.2")))
Dummy_Allow.Title.16 = data.frame(t(sapply(states52, Reg_allow, type = "AllowRate.Title.16")))
Dummy_Allow.Concurrent = data.frame(t(sapply(states52, Reg_allow, type = "AllowRate.Concurrent")))

Dummy_Allow.Total$State = rownames(Dummy_Allow.Total)
Dummy_Allow.Title.2$State = rownames(Dummy_Allow.Title.2)
Dummy_Allow.Title.16$State = rownames(Dummy_Allow.Title.16)
Dummy_Allow.Concurrent$State = rownames(Dummy_Allow.Concurrent)

Dummy_Allow.Total = MoveFront(Dummy_Allow.Total, c("State"))
Dummy_Allow.Title.2 = MoveFront(Dummy_Allow.Title.2, c("State"))
Dummy_Allow.Title.16 = MoveFront(Dummy_Allow.Title.16, c("State"))
Dummy_Allow.Concurrent = MoveFront(Dummy_Allow.Concurrent, c("State"))

#save(Dummy.Total, Dummy.Title.2, Dummy.Title.16, Dummy.Concurrent)


# Look at the R2 of allowance rates

Dummy_Allow.Total %>% arrange(desc(R2)) %>% subset(select = c("State", "R2"))
Dummy_Allow.Title.2 %>% arrange(desc(R2)) %>% subset(select = c("State", "R2"))
Dummy_Allow.Title.16 %>% arrange(desc(R2)) %>% subset(select = c("State", "R2"))
Dummy_Allow.Concurrent %>% arrange(desc(R2)) %>% subset(select = c("State", "R2"))

Allow_R2 <- 
cbind(
Dummy_Allow.Total %>% subset(select = c("State", "R2")),
Dummy_Allow.Title.2 %>% subset(select = c("R2")),
Dummy_Allow.Title.16  %>% subset(select = c("R2")),
Dummy_Allow.Concurrent %>% subset(select = c("R2"))
)
colnames(Allow_R2) <- c("State","Total", "DI", "SSI", "Concurrent")
Allow_R2 %>% arrange(desc(Total))

# R2 at national level is much higher than R2s at state level
# At national level, DI has the highest R2 while SSI has the lowest
# At state level, the pattern of R2 across the 3 types is not clear.



# Look at the correlation between R2 of application and R2 of allowance rate
which(Dummy_Allow.Total[,c("State")] == "AG")

plot(Dummy_Allow.Total[-35, "R2"], Dummy.Total[-35, "R2"])
plot(Dummy_Allow.Title.2[-35, "R2"], Dummy.Title.2[-35, "R2"])
plot(Dummy_Allow.Title.16[-35,"R2"], Dummy.Title.16[-35, "R2"])
plot(Dummy_Allow.Concurrent[-35,"R2"], Dummy.Concurrent[-35, "R2"])

cor(Dummy_Allow.Total[-35,"R2"], Dummy.Total[-35, "R2"]) # 0.4247
cor(Dummy_Allow.Title.2[-35,"R2"], Dummy.Title.2[-35, "R2"]) # 0.5077
cor(Dummy_Allow.Title.16[-35,"R2"], Dummy.Title.2[-35, "R2"]) # 0.2778
cor(Dummy_Allow.Concurrent[-35,"R2"], Dummy.Concurrent[-35, "R2"]) # 0.3538
# DI has the hightest correlation while SSI has the lowest correlation.
# But the correlation for all types is by no means strong. 


```

```{r State_by_State_Comparison_plots}
# Comparing the seasonal dummies of application and allowance rate.

# Figure 5

plot_app <- function(data, i, type){
  plot(as.numeric(subset(data, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = paste(i, type),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)
}

plot_allow <- function(data, i, type){
  plot(as.numeric(subset(data, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = paste(i, type),
       ylim = c(-0.02, 0.02),las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)
}

plot_double <- function(data1, data2, i, type, title = TRUE){
  # Application and Allowance rate: contemporary
  
#   data1 <- Dummy.Total
#   data2 <- Dummy_Allow.Total
#   type <- "Total"

 par(mar = c(3, 4, 4, 4) + 0.3)  
  
 plot(as.numeric(subset(data1, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, 
       main = ifelse(title, paste(i, type), ""),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)

 par(new = TRUE)

 plot(as.numeric(subset(data2, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, lty =3, col = "red",
       ylim = c(-0.02, 0.02), axes = FALSE)
 axis(4, las = 1, col = "red")
 legend("bottomleft", c("Application","Allowance Rate"), col = c("black", "red"), lty = c(1,3))
 
}   # allowance rate no lag

plot_double2 <- function(data1, data2, i, type, title = TRUE){
  # application and 4th lag of allowance rate
  
 #   data1 <- Dummy.Total
 #   data2 <- Dummy_Allow.Total
 #   type <- "Total"
  
  par(mar = c(3, 4, 4, 4) + 0.3)  

 plot(as.numeric(subset(data1, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = ifelse(title, paste(i, type), ""),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)

 par(new = TRUE)

 plot(as.numeric(subset(data2, State == i, select = c(paste0("month", 5:12), paste0("month", 1:4)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, lty =3, col = "red",
       ylim = c(-0.02, 0.02), axes = FALSE)
 axis(4, las = 1, col = "red")
 legend("bottomleft", c("Application","Allowance Rate 4th lead"), col = c("black", "red"), lty = c(1,3))
 
}  # 4th lag of allowance


cairo_pdf(paste0(Path, "Allow_App_Dummies.pdf"), width = 16, height = 11, onefile = TRUE)
# plot application and allowance rate(no lag) in seperate graphs
par(mfrow = c(2,4))
for(i in c(top10, states50[!states50 %in% top10])){
# i = "TX"
plot_app(Dummy.Total, i, "Total App")
plot_app(Dummy.Title.2, i, "DI App")
plot_app(Dummy.Title.16, i, "SSI App")
plot_app(Dummy.Concurrent, i, "Concurrent App")

plot_allow(Dummy_Allow.Total, i, "Total Allowance")
plot_allow(Dummy_Allow.Title.2, i, "DI Allowance")
plot_allow(Dummy_Allow.Title.16, i, "SSI Allowance")
plot_allow(Dummy_Allow.Concurrent, i, "Concurrent Allowance")
}
par(mfrow = c(1,1))
dev.off()


cairo_pdf(paste0(Path, "Allow_App_Dummies2.pdf"), width = 16, height = 11, onefile = TRUE)
par(mfrow = c(2,2))
# plot application and allowance(no lag) rate in the same graph 

for(i in c("AG", top10, states50[!states50 %in% top10])){
# i = "TX"
  
plot_double(Dummy.Total, Dummy_Allow.Total, i, "Total")
plot_double(Dummy.Title.2, Dummy_Allow.Title.2, i, "DI")
plot_double(Dummy.Title.16, Dummy_Allow.Title.16, i, "SSI")
plot_double(Dummy.Concurrent, Dummy_Allow.Concurrent, i, "Concurrent")
}
par(mfrow = c(1,1))
dev.off()


cairo_pdf(paste0(Path, "Allow_App_Dummies_4lag.pdf"), width = 16, height = 11, onefile = TRUE)
par(mfrow = c(2,2))
# plot application and allowance rate (4th lag) in the same graph 

for(i in c("AG", top10, states50[!states50 %in% top10])){
# i = "TX"
  
plot_double2(Dummy.Total, Dummy_Allow.Total, i, "Total")
plot_double2(Dummy.Title.2, Dummy_Allow.Title.2, i, "DI")
plot_double2(Dummy.Title.16, Dummy_Allow.Title.16, i, "SSI")
plot_double2(Dummy.Concurrent, Dummy_Allow.Concurrent, i, "Concurrent")
}
par(mfrow = c(1,1))
dev.off()

#### Allowance plots for the paper

# Figure 6
# 1. National level, application against allowance rate of the same month, 3 categories
cairo_pdf(paste0(Path, "Allow_App_Dummies_AG.pdf"), width = 8, height = 12, onefile = TRUE)
par(mfrow = c(3,1))
plot_double(Dummy.Title.2, Dummy_Allow.Title.2, "AG", "DI", F); title(main = "Seasonal Pattern of National Level SSDI Application and Allowance Rate")
plot_double(Dummy.Title.16, Dummy_Allow.Title.16, "AG", "SSI", F); title(main = "Seasonal Pattern of National Level SSI Application and Allowance Rate")
plot_double(Dummy.Concurrent, Dummy_Allow.Concurrent, "AG", "Concurrent", F); title(main = "Seasonal Pattern of National Level Concurent Application and Allowance Rate")

par(mfrow = c(1,1))
dev.off()

# 2. National level, application against allowance rate(4 month leads), 3 categories
cairo_pdf(paste0(Path, "Allow_App_Dummies4lead_AG.pdf"), width = 8, height = 12, onefile = TRUE)
par(mfrow = c(3,1))
plot_double2(Dummy.Title.2, Dummy_Allow.Title.2, "AG", "DI", F); title(main = "Seasonal Pattern of National Level SSDI Application and Allowance Rate")
plot_double2(Dummy.Title.16, Dummy_Allow.Title.16, "AG", "SSI", F); title(main = "Seasonal Pattern of National Level SSI Application and Allowance Rate")
plot_double2(Dummy.Concurrent, Dummy_Allow.Concurrent, "AG", "Concurrent", F); title(main = "Seasonal Pattern of National Level Concurent Application and Allowance Rate")

par(mfrow = c(1,1))
dev.off()



```


## Seasonal Pattern in number of allowance

The number of allowance, at both national and state level, show very similar seasonal pattern to the pattern in application. 
The similarity between them are greater than that between application and allowance rate. 


```{r seasonality_of_Allowance}

Path 

# Seasonal plots of national level allowance rates
select("Allow.Total", "AG", Panel) %>% monthplot
select("Allow.Title.2", "AG", Panel) %>% monthplot
select("Allow.Title.16", "AG", Panel) %>% monthplot
select("Allow.Concurrent", "AG", Panel) %>% monthplot


cairo_pdf(paste0(Path, "/AllowanceN_Monthplot.pdf"), width = 10, height = 8, onefile = T)
 par(mfrow = c(2,2))
# Seasonal plots of first differenced allowance rates
select("Allow.Total", "AG", Panel) %>% diff %>% monthplot(main = "dTotal")
select("Allow.Title.2", "AG", Panel) %>% diff %>% monthplot(main = "dTitle.2")
select("Allow.Title.16", "AG", Panel) %>% diff %>% monthplot(main = "dTitle.16")
select("Allow.Concurrent", "AG", Panel) %>% diff %>% monthplot(main = "dConcurrent")
dev.off()


# Estimated Spectral plots of first differenced allowance rates
cairo_pdf(paste0(Path, "/AllowanceN_Spectrum.pdf"), width = 10, height = 8, onefile = T)
 par(mfrow = c(2,2))
select("Allow.Total", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Total")
select("Allow.Title.2", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Title.2")
select("Allow.Title.16", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Title.16")
select("Allow.Concurrent", "AG", Panel) %>% diff %>% spec.pgram(span = c(4,4), main = "Concurrent")
dev.off()

i <- "CA"

select("Allow.Total", i, Panel)     %>% monthplot
select("Allow.Title.2", i, Panel)   %>% monthplot
select("Allow.Title.16", i, Panel)  %>% monthplot
select("Allow.Concurrent", i, Panel)%>% monthplot

select("Allow.Total", i, Panel) %>% diff %>% monthplot
select("Allow.Title.2", i, Panel) %>% diff %>% monthplot
select("Allow.Title.16", i, Panel) %>% diff %>% monthplot
select("Allow.Concurrent", i, Panel) %>% diff %>% monthplot


```

```{r Estimating_Seasonal_dummies_Allow}

variables = c("Allow.Total","Allow.Title.2", "Allow.Title.16", "Allow.Concurrent")

select("Allow.Total", "AG", Panel) %>% diff
# x <- seasonDummy(select("Allow.Total", "CA", Panel) %>% diff); x$test[,"Pr(>|t|)"]



dummyNW_AllowN = t(sapply(variables, function(x){ m = seasonDummy(select(x, "AG", Panel) %>% log %>% diff)
                                                 p.value <- m$test[,"Pr(>|t|)"]; names(p.value) <- paste0(names(p.value), ".p")
                                                 c(R2 = m$R2, m$model$coef, p.value)}))

Reg_allowN <- function(x, type ){
  m <- seasonDummy(select(type, x, Panel) %>% log %>%  diff)
  R2 <- m$R2
  Coeff <- m$model$coef
  p.value <- m$test[,"Pr(>|t|)"]
  names(p.value) <- paste0(names(p.value), ".p")
  c(R2 = R2, Coeff, p.value)
}

Dummy_AllowN.Total = data.frame(t(sapply(states52, Reg_allowN, type = "Allow.Total" )))
Dummy_AllowN.Title.2 = data.frame(t(sapply(states52, Reg_allowN, type = "Allow.Title.2")))
Dummy_AllowN.Title.16 = data.frame(t(sapply(states52, Reg_allowN, type = "Allow.Title.16")))
Dummy_AllowN.Concurrent = data.frame(t(sapply(states52, Reg_allowN, type = "Allow.Concurrent")))

Dummy_AllowN.Total$State = rownames(Dummy_AllowN.Total)
Dummy_AllowN.Title.2$State = rownames(Dummy_AllowN.Title.2)
Dummy_AllowN.Title.16$State = rownames(Dummy_AllowN.Title.16)
Dummy_AllowN.Concurrent$State = rownames(Dummy_AllowN.Concurrent)

Dummy_AllowN.Total = MoveFront(Dummy_AllowN.Total, c("State"))
Dummy_AllowN.Title.2 = MoveFront(Dummy_AllowN.Title.2, c("State"))
Dummy_AllowN.Title.16 = MoveFront(Dummy_AllowN.Title.16, c("State"))
Dummy_AllowN.Concurrent = MoveFront(Dummy_AllowN.Concurrent, c("State"))

#save(Dummy.Total, Dummy.Title.2, Dummy.Title.16, Dummy.Concurrent)
#Dummy_Allow.Total %>% fix


```

```{r State_by_State_Comparison_plots}
# Comparing the seasonal dummies of application and allowance rate.

plot_app <- function(data, i, type){
  plot(as.numeric(subset(data, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = paste(i, type),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)
}

plot_allow <- function(data, i, type){
  plot(as.numeric(subset(data, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = paste(i, type),
       ylim = c(-0.2, 0.2),las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)
}

plot_double <- function(data1, data2, i, type, title = TRUE){
  # Application and Allowance rate: contemporary
  
#   data1 <- Dummy.Total
#   data2 <- Dummy_Allow.Total
#   type <- "Total"

 par(mar = c(3, 4, 4, 4) + 0.3)  
  
 plot(as.numeric(subset(data1, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, 
       main = ifelse(title, paste(i, type), ""),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)

 par(new = TRUE)

 plot(as.numeric(subset(data2, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, lty =3, col = "red",
       ylim = c(-0.2, 0.2), axes = FALSE)
 axis(4, las = 1, col = "red")
 legend("bottomleft", c("Application","Allowance"), col = c("black", "red"), lty = c(1,3))
 
}   # allowance rate no lag

plot_double2 <- function(data1, data2, i, type, title = TRUE){
  # application and 4th lag of allowance rate
  
 #   data1 <- Dummy.Total
 #   data2 <- Dummy_Allow.Total
 #   type <- "Total"
  
  par(mar = c(3, 4, 4, 4) + 0.3)  

 plot(as.numeric(subset(data1, State == i, select = c(paste0("month", 1:12)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, main = ifelse(title, paste(i, type), ""),
       ylim = c(-0.2, 0.2), las = 1)
  axis(1, at = 1:12)
  abline(h = 0, lty = 3)

 par(new = TRUE)

 plot(as.numeric(subset(data2, State == i, select = c(paste0("month", 5:12), paste0("month", 1:4)))), 
       type = "b", ylab = "", xlab = "", lwd = 2, lty =3, col = "red",
       ylim = c(-0.2, 0.2), axes = FALSE)
 axis(4, las = 1, col = "red")
 legend("bottomleft", c("Application","Allowance 4th lead"), col = c("black", "red"), lty = c(1,3))
 
}  # 4th lag of allowance


cairo_pdf(paste0(Path, "AllowN_App_Dummies.pdf"), width = 16, height = 11, onefile = TRUE)
# plot application and allowance rate(no lag) in seperate graphs
par(mfrow = c(2,4))
for(i in c(top10, states50[!states50 %in% top10])){
# i = "TX"
plot_app(Dummy.Total, i, "Total App")
plot_app(Dummy.Title.2, i, "DI App")
plot_app(Dummy.Title.16, i, "SSI App")
plot_app(Dummy.Concurrent, i, "Concurrent App")

plot_allow(Dummy_AllowN.Total, i, "Total Allowance")
plot_allow(Dummy_AllowN.Title.2, i, "DI Allowance")
plot_allow(Dummy_AllowN.Title.16, i, "SSI Allowance")
plot_allow(Dummy_AllowN.Concurrent, i, "Concurrent Allowance")
}
par(mfrow = c(1,1))
dev.off()




cairo_pdf(paste0(Path, "AllowN_App_Dummies2.pdf"), width = 16, height = 11, onefile = TRUE)
par(mfrow = c(2,2))
# plot application and allowance(no lag) rate in the same graph 

for(i in c("AG", top10, states50[!states50 %in% top10])){
# i = "TX"
  
plot_double(Dummy.Total, Dummy_AllowN.Total, i, "Total")
plot_double(Dummy.Title.2, Dummy_AllowN.Title.2, i, "DI")
plot_double(Dummy.Title.16, Dummy_AllowN.Title.16, i, "SSI")
plot_double(Dummy.Concurrent, Dummy_AllowN.Concurrent, i, "Concurrent")
}
par(mfrow = c(1,1))
dev.off()


cairo_pdf(paste0(Path, "AllowN_App_Dummies_4lag.pdf"), width = 16, height = 11, onefile = TRUE)
par(mfrow = c(2,2))
# plot application and allowance rate (4th lag) in the same graph 

for(i in c("AG", top10, states50[!states50 %in% top10])){
# i = "TX"
  
plot_double2(Dummy.Total, Dummy_AllowN.Total, i, "Total")
plot_double2(Dummy.Title.2, Dummy_AllowN.Title.2, i, "DI")
plot_double2(Dummy.Title.16, Dummy_AllowN.Title.16, i, "SSI")
plot_double2(Dummy.Concurrent, Dummy_AllowN.Concurrent, i, "Concurrent")
}
par(mfrow = c(1,1))
dev.off()

#### 

# 1. National level, application against allowance rate of the same month, 3 categories
cairo_pdf(paste0(Path, "AllowN_App_Dummies_AG.pdf"), width = 8, height = 12, onefile = TRUE)
par(mfrow = c(3,1))
plot_double(Dummy.Title.2, Dummy_AllowN.Title.2, "AG", "DI", F); title(main = "Seasonal Pattern of National Level SSDI Application and Allowance")
plot_double(Dummy.Title.16, Dummy_AllowN.Title.16, "AG", "SSI", F); title(main = "Seasonal Pattern of National Level SSI Application and Allowance")
plot_double(Dummy.Concurrent, Dummy_AllowN.Concurrent, "AG", "Concurrent", F); title(main = "Seasonal Pattern of National Level Concurent Application and Allowance")

par(mfrow = c(1,1))
dev.off()

# 2. National level, application against allowance rate(4 month leads), 3 categories
cairo_pdf(paste0(Path, "AllowN_App_Dummies4lead_AG.pdf"), width = 8, height = 12, onefile = TRUE)
par(mfrow = c(3,1))
plot_double2(Dummy.Title.2, Dummy_AllowN.Title.2, "AG", "DI", F); title(main = "Seasonal Pattern of National Level SSDI Application and Allowance Rate")
plot_double2(Dummy.Title.16, Dummy_AllowN.Title.16, "AG", "SSI", F); title(main = "Seasonal Pattern of National Level SSI Application and Allowance Rate")
plot_double2(Dummy.Concurrent, Dummy_AllowN.Concurrent, "AG", "Concurrent", F); title(main = "Seasonal Pattern of National Level Concurent Application and Allowance Rate")

par(mfrow = c(1,1))
dev.off()



```







# 2. Relationship between seasonal components of disability applications and unemployment

## Relatinship in Deterministic Seasonality 

```{r seasonal_IV_Function}

seasonalIV = function(region){
  Title.2   = select("dlTitle.2",   region, Panel)
  Title.16  = select("dlTitle.16",  region, Panel)
  Concurrent= select("dlConcurrent",region, Panel)
  unemploy  = select("dlunemply_U", region, Panel)
  month = cycle(Title.2)
  
  DATA = cbind(Title.2 = Title.2, Title.16 = Title.16, Concurrent = Concurrent, 
               unemploy = stats:::lag.default(unemploy, -1), month)

  iv.Title.2   = ivreg(Title.2 ~   unemploy -1 | factor(month) -1, data = DATA )
  iv.Title.16  = ivreg(Title.16 ~  unemploy -1 | factor(month) -1, data = DATA )
  iv.Concurrent= ivreg(Concurrent~ unemploy -1 | factor(month) -1, data = DATA )
  
  output = data.frame(rbind(summary(iv.Title.2)$coefficients,
                      summary(iv.Title.16)$coefficients,
                      summary(iv.Concurrent)$coefficients))
  
  output = cbind(State = region, Category = c("SSDI","SSI","Concurrent"), output)
  rownames(output) = NULL
  output
}

```

```{r seasonal_IV_Top5}

# Table 5

# Coefficient for nationawilde and top5 states
seasonalIVTop5 = rbind(
                       seasonalIV("AG"),
                       seasonalIV("CA"),
                       seasonalIV("TX"),
                       seasonalIV("FL"),
                       seasonalIV("NY"),
                       seasonalIV("PA"))

names(seasonalIVTop5) = c("State", "Category", "Estimate", "Std Error", "t", "p-value" )

seasonalIVTop5

print(xtable(seasonalIVTop5[,c("State","Category", "Estimate", "Std Error")], caption = "Relationship between deterministic seasonal components of DB application and 1 month lag of unemployment"))

```

```{r seasonal_IV_all_states}

# Coefficient for nationawilde and top5 states

seaIV_states <- alply(statesAll, 1, seasonalIV) %>% rbind.fill
names(seaIV_states) = c("State", "Category", "Estimate", "Std Error", "t", "p.value" )

library("dplyr")

# number of states with sig coefficient for each type.
seaIV_states %>% arrange(Category) %>%  group_by(Category) %>% filter(p.value <= 0.05) %>% summarise(n = n())

# No. of states with coefficients of all 3 types sig
seaIV_states %>% dplyr::select(State, Category, Estimate, p.value) %>% gather(key, value, -State, - Category) %>% 
  dcast(State ~ Category + key) %>% filter(Concurrent_p.value <= 0.05, SSDI_p.value <= 0.05, SSI_p.value <= 0.05) # 15

# Among the states above, No. of states in which the coefficient of SSDI is the highest. 
seaIV_states %>% dplyr::select(State, Category, Estimate, p.value) %>% gather(key, value, -State, - Category) %>% 
  dcast(State ~ Category + key) %>% 
  filter(Concurrent_p.value <= 0.05, SSDI_p.value <= 0.05, SSI_p.value <= 0.05) %>%
  filter(SSDI_Estimate > SSI_Estimate, SSDI_Estimate > Concurrent_Estimate) # 12
  
  
  


```


```{r R2_of_application_against_R2_of_unemployment}


Dummy_R2 <- data.frame(State = Dummy.Title.2$State,
                       SSDI  = Dummy.Title.2$R2,
                       SSI   = Dummy.Title.16$R2,
                       Concurrent   = Dummy.Concurrent$R2,
                       Unemployment = Dummy.unemploy$R2)
# Correlation matrix
cor(Dummy_R2[, -1])
 # Correlation coefficients are quite low

# Regression 
summary(lm(SSDI ~ Unemployment, Dummy_R2))
summary(lm(SSI ~ Unemployment, Dummy_R2))
summary(lm(Concurrent ~ Unemployment, Dummy_R2))
# none of the slope is significant.

# Scatter Plot
ggplot(Dummy_R2, aes(x = Unemployment, y = SSDI)) + geom_point()
ggplot(Dummy_R2, aes(x = Unemployment, y = SSI))  + geom_point()
ggplot(Dummy_R2, aes(x = Unemployment, y = Concurrent)) + geom_point()
# No obvious relationship

# Scatter Plot Top 10 states
ggplot(subset(Dummy_R2, State %in% top10), aes(x = Unemployment, y = SSDI)) + geom_point()
ggplot(subset(Dummy_R2, State %in% top10), aes(x = Unemployment, y = SSI))  + geom_point()
ggplot(subset(Dummy_R2, State %in% top10), aes(x = Unemployment, y = Concurrent)) + geom_point()
# No obvious relatioship, even slightly negative relationship. 

```

## Relationship in Stochastic Seasonality 

```{r Stochastic_Seasonality}

spec.pgram(select("dlTotal", "AG", Panel, na.rm = T), spans = c(4,4))
abline(v = 1:6, lty = 3)

spec.pgram(select("dlTitle.2", "AG", Panel, na.rm = T), spans = c(4,4))
abline(v = 1:6, lty = 3)

spec.pgram(select("dlTitle.16", "AG", Panel, na.rm = T), spans = c(4,4))
abline(v = 1:6, lty = 3)



spec.pgram(select("dlunemply_U", "AG", Panel, na.rm = T), span = c(4,4))
abline(v = 1:6, lty = 3)


res.apply = ts(seasonDummy(select("dlTitle.2", "AG", Panel, na.rm = T))$model$res, s = c(2000,11), f = 12)
spec.pgram(res.apply, span = c(4,4))
abline(v = 1:6, lty = 3)

res.unemply = ts(seasonDummy(select("dlunemply_U", "AG", Panel, na.rm = T))$model$res, s = c(2000,11), f = 12)
plot(spec.pgram(res.unemply, span = c(4,4), plot = F))
abline(v = 1:6, lty = 3)


Acf(res.apply)
Acf(res.unemply, lwd = 3)
```







This document describes the seasonal patterns in U.S. Disability applications and unemployment by graphs and tables. 

plot.Seasonal.R: Comparing Seasonal dummies of applications and unemployment

plot.Heatmap.R: neater version of boxplot and heatmap of state series. 

plot.paper graphs.R

seasonalDummy.Rmd
seasonalDummyPooled.Rmd
seasonalIV.Rmd


